Structuring College Access: The Market Segment Model and College Board Geomarkets

### 2025-04-25 text/back_end_KS.qmd
rq2 LA results

>>>>

To examine how the socioeconomic and racial composition of included versus excluded prospects varies when student list purchases filter on particular Geomarkets in the Los Angeles metropolitan areas, we analyze six orders placed by a research university filtered for prospects across the entire state of California, in the 2019-2020 high school graduating class, and by PSAT scores. Two orders indicated PSAT score thresholds ranging from 1070-1180, another two orders filtered for a 1190-1260 PSAT range, and the remaining two orders filtered for scores ranging from 1270-1520.

The racial/ethnic composition of purchased student profiles within Geomarkets in the Los Angeles metropolitan area is presented in @fig-la-rq2-race. The more than 15,000 prospects in the resulting student lists at the low PSAT score thresholds (1070-1180) reflect very similar racial/ethnic compositions within Geomarkets than overall demographic makeup of the metropolitan area in 2020 (see @fig-la-rq1-race). However, the racial/ethnic composition of prospects become less proportional to overall metropolitan patterns by Geomarket as PSAT scores increase. For instance, prospect pools at middle range PSAT scores (1190-1260) within Geomarkets become disproportionately more Asian, while the representation of Hispanic and Black prospects decline substantially. This pattern becomes most pronounced at the highest PSAT score prospect pools. This disproportionate over representation for Asian prospects in comparison to the population of residents with Geomarkets ranges from 3 percentage points in the West Los Angeles and West Beach Geomarket (16% prospects versus 13% Asian residents) to 52 percentage points in the East Los Angeles Geomarket (79% prospects versus 27% residents). @fig-la-rq2-race illustrates the share of Black prospects within these high PSAT orders declines to less than 3% within all Geomarkets in comparison to Black residents making up a range of 3% (Covina and West Covina) to 18% (South and South Central Los Angeles) of residents within Geomarkets. Similar underrepresentation patterns are evident for Hispanic porspects are evident across all Geomarkets execept for a nearly equal representation in South and South Central Los Angeles (73% prospects versus 71% Hispanic residents).
